Anomalous waves triggered by abrupt depth changes: experiments and truncated KdV theory

### 2019-10-15 paper2_JNLS/paper.tex
NM

The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
\begin{align}
\label{HamStruct}
\partial_t{u} = \sympJ \vard{\Hupdn}{u}
\end{align}
where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which will take different forms on either side of the ADC. It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
\begin{align}
\label{H3H2}
\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
\end{align}
Then the Hamiltonian can be expressed as
\begin{equation}
\label{Hamiltonian}
\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree
\end{equation}
where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw} gives the separate upstream and downstream Hamiltonians as
\begin{align}
&\Hup = C_2 \, \Htwo - C_3 \, \Hthree 						&& \text{for } t<0 \\
&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree	&& \text{for } t>0
\end{align}
Here, we have chosen the sign of Lax \cite{lax1975periodic}. More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite. Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent. With their sign convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}. We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.

@@

We introduce two important invariants of KdV, the momentum and the energy
\begin{align}
\label{MomEn}
\Mo[u] \equiv \int_{-\pi}^{\pi} u \dx \, = 0 , \qquad
\En[u] \equiv \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
\end{align}
As indicated above, the momentum of $u$ vanishes since it is measured as displacement from equilibrium. Second, by the scale chosen for $\ampscale$ in \eqref{scales}, the energy has been normalized to unity.

The evolution of any functional $\mathcal{F}[u]$ is given by 
\begin{align}
\partial_t \mathcal{F} = \{ \mathcal{F}, \Ham \} = \int_{-\pi}^{\pi} \vard{\mathcal{F}}{u} \sympJ \vard{\Ham}{u} \dx
\end{align}
where again $\Ham = \Hup$ if $t<0$, and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.

@@

Briefly, consider the parameter $\lamfac$, the number of characteristic wavelengths in the physical domain. We require $\lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$, corresponding to the characteristic wavelength $\lam$ in the experiments, is resolved in the truncated dynamical system. If $\lamfac = \Lambda$, then $\lam$ corresponds to the smallest resolved wavelength. In contrast, if $\lamfac$ is chosen as an intermediate value, e.g.~$\lamfac = \Lambda/2$ or $\Lambda/4$, then the truncated system \eqref{TKdV} resolves scales that are both bigger and smaller than the characteristic value $\lam$.

Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}. The piecewise defined Hamiltonian for TKdV is given by
\begin{equation}
\label{TruncHamiltonian}
\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
\end{equation}
where $\Hthree$ and $\Htwo$ the defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
Then TKdV \eqref{TKdV} can be expressed as
\begin{align}
\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
\end{align}
where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.

The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalization values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity implies
\begin{equation}
\En[\uL] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
\end{equation}
% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
Thus, the dynamics of interest are confined to the unit hypersphere of $\En = 1$ in $\CC^{\Lambda}$.
In summary, the TKdV system possesses three important invariants: momentum, energy, and hamiltonian. The untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.

@@

	Recall that the abrupt depth change is met by traveling waves at dimensionless time $T_{ADC} = 0$, set to zero for convenience. The KdV equations describe wave dynamics over {\em long} timescales, physically $t \gg \timescale$, capturing weakly nonlinear and weakly dispersive effects. Meanwhile, evolution over shorter timescales is simply described by linear theory. The event of a wave crossing the ADC is precisely such a short-time event, and so we will employ matching conditions at the ADC that result from linear theory.

@@

	This statistical matching condition imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$. In particular, we view $\thup$ as given by the random-state of the incoming wave field. Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
\begin{equation}
\thdn = \transf \left( \thup \right)
\qquad \mbox{\em Transfer function}
\end{equation}
The {\em transfer function} $\transf$ will be a key link needed to relate the theory to experiments.

@@

	We note that the experiments discussed here preceded the development of the theory, and hence had no intent of mimicking a Gibbs measure in the wave forcing. We expect that if this forcing were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$. We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental measurements in light of the new theoretical developments.

@@

	Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $(\Lambda, \lamfac) = (16, 8)$. In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher. Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--20 (?).

@@

	Next, we aim to make the same comparison for free-surface slopes. We note that the derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.  
We therefore convert the theoretical calculated slopes values back to physical slopes via
\begin{equation}
\pd{\eta}{x} = 2 \pi^{3/2} \frac{\etastd}{\lamupdn} \pd{\uupdn}{\tilde{x}}
%\frac{\ampscale \lamfac}{\lengthscale} \pd{\uupdn}{\tilde{x}}
\end{equation}
So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode ($k = 1$),  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.

@@

	Figure \ref{slopehist} shows the distributions of surface slope that result from experiments (top) and theory (bottom). These histograms show intriguing similarities and differences. First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale. We note that, while the standard deviation of displacement was input into the TKdV theory, the slope standard deviation was not. Downstream, both the experimental and theoretical slope distributions  spread significantly. Interestingly, in the experiments, it is not the standard deviation that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but instead the excess kurtosis, which jumps from a value of 0.46 upstream to 3.4 downstream. Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50. Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant. These elevated levels of kurtosis give the distributions the look of being stretched in the tails and pointy near the centers (especially in the experiments).

@@

	Visual differences between the upstream and downstream dynamics are apparent in Fig.~\ref{trajectories}. The upstream trajectory shows several waves of modest amplitude all propagating leftward. Here, the magnitude of the positive and negative displacements are comparable. This is in stark contrast to the downstream dynamics, which feature fewer waves of larger amplitude propagating in either direction. In particular, one large wave is seen to propagate from the lower-left corner of the figure to the upper right. A bias towards positive surface displacement is visually apparent in these downstream dynamics, without even viewing the histograms.

@@

	We remind the reader that the KdV framework tracks the long-time evolution waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory. Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments, which take a fixed reference frame, without an appropriate Galilean transformation. Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed that is near $c$. The directions and speeds calculated by the TKdV framework simply determines how much the actually speed of waves deviates from $c$.

	Finally, Fig.~\ref{trajectories} sheds some light on the timescales required for wave evolution in the TKdV framework. It appears that the normalized time of $t=10$ is roughly the correct timescale to observe substantial wave dynamics. More precisely, both the left-going waves in Fig.~\ref{trajectories}(a) and the right-going waves in Fig.~\ref{trajectories}(b) require a dimensionless time of about $t=7$ to traverse the entire periodic domain. For the representative experiments, the characteristic timescale, $\timescale$, is 2.8~sec upstream and 5.7~sec downstream. Hence, a dimensionless time of $t=7$ corresponds to roughly 20~sec and 40~sec upstream and downstream respectively. Given the wave speeds ($c = $ 110 cm/s upstream and 54 cm/s downstream), these timescales can be converted to distances traveled by the waves. In fact, a simple calculation using the definitions in Section \ref{theory} gives the distance traveled as $7 \lamfac g/(2 \pi \freqp^2)$. Notice the depth dependence has serendipitously canceled. Hence, the distance traveled by the waves over a dimensionless time of $t=7$ is the same value upstream and downstream,  roughly 22 meters. Note that this distance is significantly greater than the 6-meter length of the experimental wave tank.

@@

	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how is it that the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC? The results shown in Fig.~\ref{skew-kurt} help resolve this question. This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}. The bold black curve shows the ensemble mean (ensemble size of 1000), while the skewness and kurtosis of 100 individual trajectories are shown by the faint gray curves to give a sense for the variance involved. Importantly, the skewness and kurtosis evolve on a much shorter timescale than the aforementioned $t=7$. More precisely, skewness and kurtosis have already saturated to their asymptotic by $t = 1$, and reach half of that value by $t = 0.18$. These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments. Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter the statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.

@@

This section discusses perhaps the most important single result of the manuscript. In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$, in particular how that ratio depends on driving amplitude and depth ratio. Below, we recap the derivation of this formula and then test the prediction against our direct experimental measurements.

>>>>

The variable-depth KdV \eqref{dimlessKdV}, though not Hamiltonian throughout the entire domain, admits a Hamiltonian structure on each side of the ADC. Indeed, \eqref{dimlessKdV} can be expressed as
\begin{align}
\label{HamStruct}
\partial_t{u} = \sympJ \vard{\Hupdn}{u}
\end{align}
where $\sympJ = \pdi{}{x}$ is the symplectic operator and $\Ham = \Hupdn$ is the Hamiltonian, which will take different forms on either side of the ADC. It is convenient to decompose the Hamiltonian into a so-called cubic and quadratic component, given respectively by
\begin{align}
\label{H3H2}
\Hthree = \frac{1}{6} \int_{-\pi}^{\pi} u^3 \dx	\, , \qquad
\Htwo = \frac{1}{2} \int_{-\pi}^{\pi} u_x^2 \dx	\, .
\end{align}
Then the Hamiltonian can be expressed as
\begin{equation}
\label{Hamiltonian}
\Hupdn = C_2 \dratupdn^{1/2} \, \Htwo - C_3 \dratupdn^{-3/2} \, \Hthree
\end{equation}
where $\drat = \dratupdn$ changes value across the ADC. More explicitly, substituting \eqref{dratpw} gives the separate upstream and downstream Hamiltonians as
\begin{align}
&\Hup = C_2 \, \Htwo - C_3 \, \Hthree 						&& \text{for } t<0 \\
&\Hdn = C_2 \dratdn^{1/2} \, \Htwo - C_3 \dratdn^{-3/2} \, \Hthree	&& \text{for } t>0
\end{align}
Here, we have chosen the sign convention of Lax \cite{lax1975periodic}. More recent work of Bajars et al.~\cite{bajars2013weakly} and Majda et al.~\cite{majda2019} use a different convention, in which both the signs of $\sympJ$ and $\Ham$ are opposite. Clearly, these sign differences cancel in \eqref{HamStruct} and thus the two conventions are completely equivalent. Using the second convention, Majda et al.~found that a {\em negative} inverse temperature is required to accurately describe the experimental observations \cite{majda2019}. We have  chosen the convention above so that a {\em positive} inverse temperature may be used, allowing our theory to fit into the most standard form of statistical mechanics.

@@

We introduce two important invariants of KdV, namely the momentum and the energy
\begin{align}
\label{MomEn}
\Mo[u] \equiv \int_{-\pi}^{\pi} u \dx \, = 0 , \qquad
\En[u] \equiv \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
\end{align}
As indicated above, the momentum of $u$ vanishes since it is measured as displacement from equilibrium. Second, due to the choice of $\ampscale$ in \eqref{scales}, the energy has been normalized to unity.

% UNUSED POISSON BRACKET STUFF
\begin{comment}
The evolution of any functional $\mathcal{F}[u]$ is given by 
\begin{align}
\partial_t \mathcal{F} = \{ \mathcal{F}, \Ham \} = \int_{-\pi}^{\pi} \vard{\mathcal{F}}{u} \sympJ \vard{\Ham}{u} \dx
\end{align}
where again $\Ham = \Hup$ if $t<0$, and $\Ham = \Hdn$ if $t>0$ and $\{\}$ is the Poisson bracket induced by $\sympJ$.
\end{comment}

@@

Briefly, consider the parameter $\lamfac$, the number of characteristic wavelengths in the physical domain. We require $1 \le \lamfac \le \Lambda$, so that the mode $\uhat_{\lamfac}$, corresponding to the characteristic wavelength $\lam$ in the experiments, is resolved in the truncated dynamical system. If $\lamfac = \Lambda$, then $\lam$ corresponds to the smallest resolved wavelength. If instead $\lamfac$ is chosen as an intermediate value between 1 and $\Lambda$, then the truncated system will resolve scales that are both bigger and smaller than the characteristic value $\lam$.

Remarkably, the TKdV system \eqref{TKdV} retains nearly the same Hamiltonian structure described in Section \ref{HamiltonianSection}, with the only modification being the inclusion of the projection operator \cite{bajars2013weakly, majda2019}. The piecewise defined Hamiltonian for TKdV is given by
\begin{equation}
\label{TruncHamiltonian}
\HLupdn = C_2 \dratupdn^{1/2} \, \Htwo[\uL] - C_3 \dratupdn^{-3/2} \, \Hthree[\uL]
\end{equation}
where $\Hthree$ and $\Htwo$ are defined exactly as before \eqref{H3H2}, but now are simply applied to the projected variable $\uL = \Proj u$.
Then TKdV \eqref{TKdV} can be expressed as
\begin{align}
\partial_t {\uL} = \partial_x \Proj \, \vard{\HLupdn}{\uL}
\end{align}
where the truncated symplectic operator is $\SympL = \partial_x \Proj$.
% Note: Bajars 2013 has some important details on the Hamiltonian on the truncated system.

The momentum and energy defined in \eqref{MomEn} remain invariants of TKdV, with the same normalization values $\Mo[\uL] = 0$ and $\En[\uL] = 1$. Note that Parseval's identity implies
\begin{equation}
\En[\uL] = 2 \pi \sum_{k=1}^{\Lambda} \abs{\uhat_k}^2 = 1
\end{equation}
% The above agrees with the first line on top of p. 4 in Majda and Qi J Stat Phys 2019.
Thus, the dynamics of interest are confined to the unit hypersphere, $\En = 1$ in $\CC^{\Lambda}$.
In summary, the TKdV system possesses three important invariants: momentum, energy, and Hamiltonian. The untruncated KdV system possesses an infinite sequence of additional invariants \cite{lax1975periodic, whitham2011linear}, but their truncated counterparts are not generally invariants of TKdV.

@@

	Recall that the abrupt depth change is met by traveling waves at dimensionless time $T_{ADC} = 0$, set to zero for convenience. The KdV equations describe wave dynamics over {\em long} timescales, physically $t \gg \timescale$, capturing weakly nonlinear and weakly dispersive effects. Meanwhile, evolution over shorter timescales is simply described by linear theory. The event of a wave crossing the ADC is precisely such a short-time event, and so we will employ the same matching conditions at the ADC that would result from linear theory.

@@

	This statistical matching condition, originally derived by Majda et al., imposes a relationship between the two inverse temperatures $\thup$ and $\thdn$. In particular, we view $\thup$ as given by the random-state of the incoming wave field. Thus, while we will treat $\thup$ as a free parameter that can be varied to study various possible system states, the downstream $\thdn$ is determined directly by the matching condition \eqref{statmatch}, giving the functional dependence
\begin{equation}
\thdn = \transf \left( \thup \right)
\qquad \mbox{\em Transfer function}
\end{equation}
The {\em transfer function} $\transf$ will be a key link needed to relate the theory to experiments.

@@

	We note that the experiments discussed here preceded the development of the theory, and hence had no intent of mimicking a Gibbs measure in the wave forcing. We expect that if the experimental forcing were designed to mimic the Gibbs measure, in particular the spectral decay, then it would be more straightforward to assign a value to the model parameter $\lamfac$. We, however, leave that task for future research due to the significant cost of performing an entirely new set of experiments compared to the relative ease and great value in re-analyzing existing experimental measurements in light of the new theoretical developments.

@@

	Next, Fig.~\ref{transfig}(b) shows the skewness of the incoming (green) and outgoing (blue) wave fields, as they depend on the incoming inverse temperature $\thup$ for the case $(\Lambda, \lamfac) = (16, 8)$. In line with experimental observations, the incoming skewness is small, while the outgoing wave skewness is much higher. Specifically, to capture the experimentally observed peak skewness range of 0.6-0.9 seen for the larger amplitudes in Fig.~\ref{ExpSpatialStats}(b), requires selecting $\thup$ in the range 10--25.

@@

	Next, we aim to make the same comparison for free-surface slopes. We note that the derivative, $\partial \eta / \partial x$, is already a dimensionless quantity with a simple physical interpretation, namely the slope of the free surface, whereas the interpretation of $\partial u/\partial \tilde{x}$ is tied to the characteristic values $\ampscale$ and $\lengthscale$.  
We therefore convert the theoretical calculated slopes values back to physical slopes via
\begin{equation}
\pd{\eta}{x} = 2 \pi^{3/2} \frac{\etastd}{\lamupdn} \pd{\uupdn}{\tilde{x}}
%\frac{\ampscale \lamfac}{\lengthscale} \pd{\uupdn}{\tilde{x}}
\end{equation}
So that the peak experimental wavelength $\lam$ (corresponding to $\freqp = 2$ Hz) would be mapped to the dominating TKdV spectral mode of $k = 1$,  this conversion formula has not been corrected with the extra factor of $\lamfac$ present in \eqref{scales}.

@@

	Figure \ref{slopehist} shows the distributions of surface slope that result from experiments (top) and theory (bottom). These histograms show intriguing similarities and differences. First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale. We note that, while the standard deviation of displacement was input into the TKdV theory, the slope standard deviation was not. Downstream, both the experimental and theoretical slope distributions  spread significantly. Interestingly, in the experiments, it is not the standard deviation that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but instead the excess kurtosis, which jumps from a value of 0.46 upstream to 3.4 downstream. Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50. Thus, the theoretically predicted jump in kurtosis, while not as extreme as that measured in experiments, is quite significant. These elevated levels of kurtosis are associated with the distinct appearance of the distribution, having flat tails and somewhat sharp centers.

@@

	Visual differences between the upstream and downstream dynamics are apparent in Fig.~\ref{trajectories}. The upstream trajectory shows several waves of modest amplitude all propagating leftward. Here, the magnitude of the positive and negative displacements are comparable. This is in stark contrast to the downstream dynamics, which feature fewer waves of larger amplitude propagating in either direction. In particular, one large wave is seen to propagate from the lower-left corner of the figure to the upper right. A bias towards positive surface displacement is apparent in these downstream dynamics, without even viewing the histograms.

@@

	We remind the reader that the KdV framework tracks the long-time evolution of waves in a reference frame moving with the characteristic speed $c = \sqrt{g \depth}$ from linear theory. Hence, the above discussion of left-or-right going waves cannot be interpreted in the context of the experiments, which take a fixed reference frame, without an appropriate Galilean transformation. Hence, in the laboratory frame, all waves indeed propagate unidirectionally, from left to right, with a speed that is near $c$. The directions and speeds calculated by the TKdV framework simply determines how much the true wave speed deviates from $c$.

	Finally, Fig.~\ref{trajectories} sheds some light on the timescales required for wave evolution in the TKdV framework. It appears that the normalized time of $t=10$ is roughly the correct timescale to observe substantial wave dynamics. More precisely, both the left-going waves in Fig.~\ref{trajectories}(a) and the right-going waves in Fig.~\ref{trajectories}(b) require a dimensionless time of about $t=7$ to traverse the entire periodic domain. For the representative experiments, the characteristic timescale, $\timescale$, is 2.8~sec upstream and 5.7~sec downstream. Hence, a dimensionless time of $t=7$ corresponds to roughly 20~sec and 40~sec upstream and downstream respectively. Given the wave speeds ($c = $ 110 cm/s upstream and 54 cm/s downstream), these timescales can be converted to distances traveled by the waves. In fact, a simple calculation using the definitions in Section \ref{theory} gives the distance traveled as $7 \lamfac g/(2 \pi \freqp^2)$, which, due to a serendipitous cancellation, is independent of depth. Hence, the distance traveled by the waves over a dimensionless time of $t=7$ is the same value upstream and downstream,  roughly 22 meters. Note that this distance is significantly greater than the 6-meter length of the experimental wave tank.

@@

	These estimates raise an important question: if a distance of 22 meters is required for significant wave evolution under the KdV framework, how do the experiments exhibit substantial changes in wave statistics a much shorter distance downstream of the ADC? The results shown in Fig.~\ref{skew-kurt} help resolve this question. This figure shows the evolution of surface-displacement skewness (left) and kurtosis (right) computed by the same TKdV deterministic simulations as pictured in Fig.~\ref{trajectories}. The bold black curve shows the ensemble mean (ensemble size of 1000), while the faint gray curves show the skewness and kurtosis of 100 individual trajectories to give a sense for the variation involved. Importantly, the skewness and kurtosis evolve on a much shorter timescale than the aforementioned $t=7$. More precisely, skewness and kurtosis have already saturated to their asymptotic value by $t = 1$, and reach half of that value by $t = 0.18$. These dimensionless times correspond to travel distances of 310 cm and 56 cm respectively --- much shorter than the previously mentioned 22 meters, and on the same order as the relevant distances in the experiments. Thus, after crossing the ADC, the wave-field rapidly reconfigures itself enough to fundamentally alter its statistical distributions, and this reconfiguration occurs on a much shorter timescale than that required for waves to cross the entire periodic domain.

@@

We now discuss in this section arguably the most significant single result of the manuscript. In the recent theoretical study, Majda et al.~2019 derived an explicit formula for the outgoing wave-field skewness in terms of the system parameters \cite{majda2019}. More specifically, this formula relates the downstream skewness of surface displacement $\skw(\eta)$ to the change in slope variance $\var(\eta_x)$. Below, we recap the derivation of this formula and then test the prediction against our direct experimental measurements.
